---
title: "Modeling"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidymodels)
library(modelr)
library(broom)
doParallel::registerDoParallel()
```

```{r}
df_all_only_total_na <- read_csv("./data/processed/combined_total_na.csv")
df_all_only_total <- read_csv("./data/processed/combined_total.csv")

df_all_only_enteric_na <- read_csv("./data/processed/combined_enteric.csv")
df_all_only_enteric <- read_csv("./data/processed/combined_enteric.csv")
df_livestock_patterns <- read_csv("./data/processed/df_livestock_patterns.csv")
```

```{r}
countries_bad = c("USSR", "Americas", "Asia", "Australia and New Zealand", "China", "Central Asia", "Central America", "Eastern Africa", "Eastern Asia", "Eastern Euorpe", "Europe", "Land Locked Developing Countries", "Least Developed Countries", "Low Income Food Deficit Countries", "Middle Africa", "Net Food Importing Developing Countries", "Northern Europe", "Northern Africa", "Northern America", "Pacific Islands Trust Territory", "Southern Asia", "South America", "Southern Africa", "Small Island Developing States", "Serbia and Montenegro", "Western Asia", "Western Africa")
df_all_only_total <-
  df_all_only_total_na %>%
  rename(CO2 = `Agriculture total|CO2`) %>%
  # df_all_only_enteric_na %>%
  # rename(CO2 = `Enteric Fermentation|CO2`) %>%
  # filter(Area == c("United States of America")) %>%
  # select_if(~ !any(is.na(.))) %>%
  select_if(~ !is.logical(.)) %>%
  # filter(Year >= 1990) %>%
  replace(is.na(.), 0) %>%
  # select(contains("CO2"), Year, Area) %>%
  filter(CO2 != 0) %>%
  select(-contains("and")) %>%
  filter(!(Area %in% countries_bad))
  # select(contains("CO2"), Year, Area, contains("Stock"))
  # select(contains("CO2"), Year, Area, !contains("Stock"))
  # select(contains("CO2"), Year, Area, contains("Area"))
# -contains("Area harvested"), -contains("Production") contains("CO2"), Year, Area, contains("Yield") , -contains("Stocks")
# df_all_only_total <-
#   df_livestock_patterns

  df_all_only_total %>% select(contains("Stock")) %>% names() %>% as.factor() %>% fct_unique()

# df_all_only_enteric <-
  # df_all_only_enteric %>%
  # filter(Area == "United States of America")
#   df_all_only_enteric_na %>%
#   filter(Area == "United States of America") %>%
#   select_if(~ !any(is.na(.)))
# df_all_only_enteric
```


```{r}
emissions_split = initial_split(df_all_only_total, prop = 0.8)
emissions_train = training(emissions_split)
emissions_test = testing(emissions_split)
emissions_train
emissions_test
```

```{r}
emissions_rec <-
  # recipe(`Enteric Fermentation|CO2` ~ ., data = emissions_train) %>%
  recipe(`CO2` ~ ., data = emissions_train) %>%
  step_rm(Year, Area) %>%
  step_zv(all_numeric(), -all_outcomes()) %>%
  step_normalize(all_numeric(), -all_outcomes())

emissions_prep <-
  emissions_rec %>%
  prep()
```

```{r}
lasso_spec <-
  linear_reg(penalty = 0.1, mixture=1) %>%
  set_engine("glmnet")

wf <-
  workflow() %>%
  add_recipe(emissions_rec)

lasso_fit <-
  wf %>%
  add_model(lasso_spec) %>%
  fit(data = emissions_train)

df_coeff <-
  lasso_fit %>%
  pull_workflow_fit() %>%
  tidy()
df_coeff
```

```{r}
df_coeff %>%
  arrange(desc(estimate)) %>%
  filter(term != "(Intercept)")
```

```{r}
df_coeff %>% 
  mutate(
    term = fct_reorder(term, estimate)
  ) %>% 
  arrange(desc(abs(estimate))) %>% 
  filter(term != "(Intercept)") %>%
  head(30) %>%
  # filter(estimate > 0) %>%
  ggplot(aes(
      x = estimate,
      y = term
    )) +
  geom_col()
```

```{r}
emissions_boot <- bootstraps(emissions_train, 25)

tune_spec <-
  linear_reg(penalty = tune(), mixture = 1) %>%
  set_engine("glmnet")

lambda_grid <- grid_regular(penalty(), levels = 50)
```

```{r}
lasso_grid <-
  tune_grid(
    wf %>% add_model(tune_spec),
    resamples = emissions_boot,
    grid = lambda_grid
  )
```

```{r}
lasso_grid %>%
  collect_metrics() %>%
  ggplot(aes(penalty, mean, color = .metric)) +
  geom_errorbar(aes(
    ymin = mean - std_err,
    ymax = mean + std_err
  ),
  alpha = 0.5
  ) +
  geom_line(size = 1.5) +
  facet_wrap(~.metric, scales = "free", nrow = 2) +
  scale_x_log10() +
  theme(legend.position = "none")
```

```{r}
lowest_rmse <- lasso_grid %>%
  select_best("rmse")

final_lasso <- finalize_workflow(
  wf %>% add_model(tune_spec),
  lowest_rmse
)
```

```{r}
final_fit <- last_fit(
  final_lasso,
  emissions_split
)

emissions_split
final_fit %>% collect_metrics()
```

```{r}
folds <- vfold_cv(emissions_train, v = 10)

results <-
  fit_resamples(
    final_lasso,
    folds
  ) %>%
  collect_metrics()
results
```

```{r}
confidence_level = 0.95

stat <- results %>% filter(.metric == "rmse")
lo <- pull(stat, "mean") - qnorm( 1 - (1 - confidence_level) / 2 ) * pull(stat, "std_err")
hi <- pull(stat, "mean") + qnorm( 1 - (1 - confidence_level) / 2 ) * pull(stat, "std_err")
lo
hi

stat <- results %>% filter(.metric == "rsq")
lo <- pull(stat, "mean") - qnorm( 1 - (1 - confidence_level) / 2 ) * pull(stat, "std_err")
hi <- pull(stat, "mean") + qnorm( 1 - (1 - confidence_level) / 2 ) * pull(stat, "std_err")
lo
hi
```

```{r}
predictions <-
  collect_predictions(
    last_fit(
      final_lasso,
      emissions_split,
    )
  ) %>%
  pull(".pred")

testing(emissions_split) %>%
  mutate("Predicted CO2" = predictions) %>%
  # filter(Area == "United States of America") %>%
  filter(CO2 < 500000) %>%
  ggplot(aes(CO2, `Predicted CO2`)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, color = "grey")

emissions_test %>%
  mutate("Predicted CO2" = predictions) %>%
  # filter(Area == "United States of America") %>%
  filter(CO2 < 1000000) %>%
  # filter(CO2 > 100000) %>%
  group_by(Area) %>%
  filter(n() > 10) %>%
  ungroup() %>%
  pivot_longer(
    cols = c("CO2", "Predicted CO2"),
    names_to = "metric",
    values_to = "value"
  ) %>%
  ggplot(aes(x = Year, group = Area, color = Area)) +
  geom_line(aes(y = value)) +
  # scale_y_log10() +
  # scale_x_log10() +
  theme(legend.position = "none") +
  facet_wrap(~ metric)
```




 All countries, all years, only yield: 0.607
 All countries, all years, only yield, no logical: 0.587
 All countries, all years, only area, no logical: 0.977
 All countries, all years, only crops, no logical: 0.998
 All countries, all years, only livestock: 0.975
 All countries, all years, all inputs: 0.985
 All countries, filtered years, only yield: 0.684
*All countries, filtered years, only yield, no logical: 0.271*
 All countries, filtered years, only livestock: 0.979
 All countries, filtered years, all inputs: 0.999


*Maybe Causation?*
- Impact of different factors on agricultural emissions.
- How to modify current agricultural produce to produce less emissions?

- What are the best predictors for predicting agricultural emissions?
- Can we build a model that predicts agricultural emission from the raw agricultural output?
  - All countries, predicted versus actual
  - All countries, predicted versus actual, animated over time
    - Requires creating test set from last n years 

Above for enteric or total agricultural emissions
















  








